---
title: "pyscenic_import_r"
output: html_document
date: "2024-01-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Importing the results from pyscenic 
```{r}
pyScenicDir <- "./data/week30/pyscenic"
library(SCENIC)
library(tidyverse)
```

#Reading the loom file 
```{r}
library(SCopeLoomR)
pyScenicLoomFile <- file.path(pyScenicDir,"output.loom")
loom <- open_loom(pyScenicLoomFile, mode = "r")
```

#Read info from the loom file 
```{r}
# Read information from loom file:
regulons_incidMat <- get_regulons(loom, column.attr.name='Regulons')
regulons <- regulonsToGeneLists(regulons_incidMat)
regulonAUC <- get_regulons_AUC(loom, column.attr.name='RegulonsAUC')
regulonsAucThresholds <- get_regulon_thresholds(loom)
#embeddings <- get_embeddings(loom)
```



```{r}
exprMat <- get_dgem(loom)
cellInfo <- get_cell_annotation(loom)
#clusterings <- get_clusterings_with_name(loom)
```




```{r}
cellInfo <- readRDS(file = "./data/week30/rda/cellInfo.rds")
cellInfo$seurat_clusters_ind <- paste(cellInfo$genotype,cellInfo$seurat_clusters,
                                      sep = "_")
```


```{r}
library(AUCell)
library(grid)
cellInfo <- data.frame(cellInfo)
#regulonAUC <- loadInt(scenicOptions, "aucell_regulonAUC")
regulonAUC <- regulonAUC[onlyNonDuplicatedExtended(rownames(regulonAUC)),]

regulonActivity_byCellType <- sapply(split(rownames(cellInfo), cellInfo$genotype_clusters),
                                     function(cells) rowMeans(getAUC(regulonAUC)[,cells]))

regulonActivity_byCellType_Scaled <- t(scale(t(regulonActivity_byCellType), center = T, scale=T))

ComplexHeatmap::Heatmap(regulonActivity_byCellType, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 7))
#Be careful about scaling and centering the data

ComplexHeatmap::Heatmap(regulonActivity_byCellType_Scaled, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 7)
                        )
```

##Subset on cell types and compare wild type vs mutant  
```{r}
#Subsetting on macrophages 

#Fisrt subset only on the raw values form the regulon activity as scaling is done based on other cell types too 

#First only on "Inflmatory Macrophages"
columns_to_keep <- grep("Inflamatory Macrophages", colnames(regulonActivity_byCellType), value = TRUE)

# Subset the data frame to keep only the selected columns
regulonActivity_InflamatoryMacs <- regulonActivity_byCellType[, columns_to_keep]

var_InflamatoryMacs <- apply(regulonActivity_InflamatoryMacs,1,var)
# Sort rows based on variances
sorted_rows <- order(var_InflamatoryMacs, decreasing = TRUE)
# Select the top 50 rows
top_50_InflamtoryMacs <- regulonActivity_InflamatoryMacs[sorted_rows[1:50], ]

#First lets make a heatmap to get an idea about the regulons that are most likely to be changed 
#Then we can plot histograms with the distribution of their AUC activity 

#Scale now within the celltype 
top_50_InflamtoryMacs_scaled <- t(scale(t(top_50_InflamtoryMacs), center = T, scale=T))

ComplexHeatmap::Heatmap(top_50_InflamtoryMacs, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 6))

```


```{r}
#Scaled version of the heatmap 

ComplexHeatmap::Heatmap(top_50_InflamtoryMacs_scaled, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 7)
                        )
```


#We need to get the regulonActivity again : This time by individual cell types and samples : So 5 in total

```{r}
regulonActivity_byCellType_ind <- sapply(split(rownames(cellInfo), cellInfo$seurat_clusters_ind),
                                     function(cells) rowMeans(getAUC(regulonAUC)[,cells]))
```



##Subset on cell types and compare wild type vs mutant(Differnece being we are doing on an individual sample scale)
```{r}
#Subsetting on macrophages 

#Fisrt subset only on the raw values form the regulon activity as scaling is done based on other cell types too 

#First only on "Inflmatory Macrophages"
columns_to_keep <- grep("Inflamatory Macrophages", colnames(regulonActivity_byCellType_ind), value = TRUE)

# Subset the data frame to keep only the selected columns
regulonActivity_InflamatoryMacs_ind <- regulonActivity_byCellType_ind[, columns_to_keep]

var_InflamatoryMacs_ind <- apply(regulonActivity_InflamatoryMacs_ind,1,var)
# Sort rows based on variances
sorted_rows <- order(var_InflamatoryMacs_ind, decreasing = TRUE)
# Select the top 50 rows
top_50_InflamtoryMacs_ind <- regulonActivity_InflamatoryMacs_ind[sorted_rows[1:50], ]

#First lets make a heatmap to get an idea about the regulons that are most likely to be changed 
#Then we can plot histograms with the distribution of their AUC activity 

#Scale now within the celltype 
top_50_InflamtoryMacs_scaled_ind <- t(scale(t(top_50_InflamtoryMacs_ind), center = T, scale=T))

ComplexHeatmap::Heatmap(top_50_InflamtoryMacs_ind, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 6))

```


```{r}
#Scaled version of the heatmap 

ComplexHeatmap::Heatmap(top_50_InflamtoryMacs_scaled_ind, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 7)
                        )
```

##Do the exact same thing on Resident-like Macrophages 

```{r}
#Now on "Resident-like Macrophages"
columns_to_keep <- grep("Resident-like Macrophages", colnames(regulonActivity_byCellType), value = TRUE)

# Subset the data frame to keep only the selected columns
regulonActivity_resMacs <- regulonActivity_byCellType[, columns_to_keep]

var_resMacs <- apply(regulonActivity_resMacs,1,var)
# Sort rows based on variances
sorted_rows <- order(var_resMacs, decreasing = TRUE)
# Select the top 50 rows
top_50_resMacs <- regulonActivity_resMacs[sorted_rows[1:50], ]

#First lets make a heatmap to get an idea about the regulons that are most likely to be changed 
#Then we can plot histograms with the distribution of their AUC activity 

#Scale now within the celltype 
top_50_resMacs_scaled <- t(scale(t(top_50_resMacs), center = T, scale=T))

ComplexHeatmap::Heatmap(top_50_resMacs, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 6))

```


```{r}
#Scaled version of the heatmap for resident-like macrophages

ComplexHeatmap::Heatmap(top_50_resMacs_scaled, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 7)
                        )
```

##Resident like macrophages at the individual level 

```{r}

#on "Resident-like Macrophages"
columns_to_keep <- grep("Resident-like Macrophages", colnames(regulonActivity_byCellType_ind), value = TRUE)

# Subset the data frame to keep only the selected columns
regulonActivity_resMacs_ind <- regulonActivity_byCellType_ind[, columns_to_keep]

var_resMacs_ind <- apply(regulonActivity_resMacs_ind,1,var)
# Sort rows based on variances
sorted_rows <- order(var_resMacs_ind, decreasing = TRUE)
# Select the top 50 rows
top_50_resMacs_ind <- regulonActivity_resMacs_ind[sorted_rows[1:50], ]

#First lets make a heatmap to get an idea about the regulons that are most likely to be changed 
#Then we can plot histograms with the distribution of their AUC activity 

#Scale now within the celltype 
top_50_resMacs_scaled_ind <- t(scale(t(top_50_resMacs_ind), center = T, scale=T))

ComplexHeatmap::Heatmap(top_50_resMacs_ind, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 6))

```



#Scaled version for resident like macs 

```{r}
#Scaled version of the heatmap 

ComplexHeatmap::Heatmap(top_50_resMacs_scaled_ind, 
                        name="Regulon activity",
                        row_names_gp = grid::gpar(fontsize = 8),
                        column_names_gp = grid::gpar(fontsize = 7)
                        )
```

#Write a function to get the AUC values, plot the heatmap for ind vs all together, so that dont have to copy paste so many times 




##Plot histogram of AUC values for a few tfs 
```{r}
#First step is to get the regulon activity of individual cells 
regulonActivity_byCellType <- sapply(split(rownames(cellInfo), cellInfo$genotype_clusters),
                                     function(cells) rowMeans(getAUC(regulonAUC)[,cells]))

#Have this huge matrix with all cells vs the 266 TFs at an individual level 
regulonActivity_by_indcells <- getAUC((regulonAUC))

#Subset for resident macrophages 
cellInfo_resMacs <- cellInfo[cellInfo$seurat_clusters == "Resident-like Macrophages",]
selected_cols <- intersect(colnames(regulonActivity_by_indcells),rownames(cellInfo_resMacs))
regulonActivity_resMacs_ind_cells <- regulonActivity_by_indcells[,selected_cols]

#Can select tf for example : Here we select Jun 
regulonActivity_by_indcells_Jun <- regulonActivity_resMacs_ind_cells %>%
                                   t() %>%
                                   as.data.frame() %>%
                                   select("Jun(+)") %>%
                                   rename(Jun = "Jun(+)")

combined_ann_Jun <- merge(as.data.frame(regulonActivity_by_indcells_Jun), 
                          cellInfo_resMacs, by = "row.names", 
                          all.x = TRUE)
library(ggplot2)


ggplot(combined_ann_Jun, aes(x = Jun, fill = genotype)) +
  geom_histogram(aes(y = ..density..), alpha = 0.5, position = "identity", bins = 30) +
  geom_density(alpha = 0.6) +
  labs(title = "Histogram and Density Plot by Sample", x = "Jun(+)", y = "Density") +
  theme_minimal()
  
     
```




```{r}
ggplot(combined_ann_Jun, aes(x = Jun, fill = mutation)) +
  geom_histogram(aes(y = ..density..), alpha = 0.5, position = "identity", bins = 30) +
  geom_density(alpha = 0.5) +
  labs(title = "Histogram and Density Plot by Genotype", x = "Jun(+)", y = "Density") +
  theme_minimal()+
  scale_fill_viridis_d()
```







##Have to find a way to set the thresholds for binarized version of the regulon activity 

```{r}

```



##Find top regulators 
```{r}

```














```{r}
cellInfo <- data.frame(cellInfo)
regulonAUC <- loadInt(scenicOptions, "aucell_regulonAUC")
regulonAUC <- regulonAUC[onlyNonDuplicatedExtended(rownames(regulonAUC)),]
regulonActivity_byCellType <- sapply(split(rownames(cellInfo), cellInfo$genotype_clusters),
                                     function(cells) rowMeans(getAUC(regulonAUC)[,cells]))
regulonActivity_byCellType_Scaled <- t(scale(t(regulonActivity_byCellType), center = T, scale=T))

ComplexHeatmap::Heatmap(regulonActivity_byCellType, name="Regulon activity")
#Be careful about scaling and centering the data
ComplexHeatmap::Heatmap(regulonActivity_byCellType_Scaled, name="Regulon activity")
```










